Supplementary Information

### 2015-09-02 manuscript/latitudepaper.tex
DBS comments

\title{Latitudinal gradients in niche breadth vary across ecosystem types.}
\author{Alyssa R. Cirtwill$^{1,2}$, Daniel B. Stouffer$^{1}$, Tamara N. Romanuk$^{2}$}
\date{\small$^1$Centre for Integrative Ecology\\School of Biological Sciences\\University of Canterbury\\
Private Bag 4800\\Christchurch 8140, New Zealand \\
\medskip$^2$Department of Biology\\
Life Science Centre, Dalhousie University\\1355 Oxford St., P0 BOX 15000\\
Halifax NS, B3H 4R2, Canada\\}

@@

  If the latitude-niche breadth hypothesis is correct, there should also be direct relationships between latitude and the
  degree of specialisation (i.e., the breadth of the Eltonian niche; ~\cite{Elton1927,Leibold2010}) of species within food webs.
  Specifically, narrower niches in the tropics would equate to
  greater specialisation (narrower niches) while constant niche sizes but
  greater productivity would translate to constant specialisation and niche
  width across latitude.
  Attempts to unravel these effects, however,
  are complicated by known relationships between species richness and many other network properties~\cite{Riede2010}. 
  For example,
  narrower niches imply fewer links per species (i.e., greater specialisation) in the tropics (\cite{Marra1997,Dyer2007};
  but see~\cite{Schleuning2012}). However,
  average numbers of links per species tend to increase in larger food webs~\cite{Dunne2006,Riede2010}. This means
  that a latitudinal effect on specialisation may be obscured by a latitudinal gradient in species richness.

  If this is the case, it may still be possible to uncover effects of latitude
  on specialisation by examining the shape of the scaling  relationship
  between specialisation and species richness over changing latitude. As
  food webs describing different ecosystem types may differ both in their
  food-web structure~\cite{Dunne2004,Shurin2006} and in the extent to [[seasonal
  variation? Latitudinal gradients in S? Productivity? whatnow?]], we also included
  ecosystem type as a covariate.   Here, we
  use three measures of specialisation: mean links per species, mean
  generality (number of prey), and mean vulnerability (number of predators).
  By testing whether latitude affects the scaling of each property with
  species richness, we test for the effects of latitude on specialisation
  implied by the ``latitude-niche breadth hypothesis''. If the scaling of
  specialisation with species richness is weaker in the tropics (i.e., if
  species gain fewer links, predators, or prey as the size of the network
  increases), this will indicate narrower niches at the tropics. If, however,
  the scaling of specialisation with species richness does not vary over
  latitude, this will indicate that niches are similarly-sized
  worldwide but that there is a broader niche space in the tropics.

@@

  \subsection*{Gradients over latitude}

@@

    \noindent where $S_{i}$ is the species richness of web $i$, $L_{i}$ its absolute
    latitude (degrees north or south  regardless of direction), $E_{i}$ is a categorical
    variable indicating the ecosystem type of network $i$ (comprising terms for stream, 
    marine, lake, and terrestrial networks with estuarine
    networks providing the intercept) and $\epsilon_{i}$ is a residual error term. 
    We next calculated the AIC
    of the maximal model as well as the AIC's of a suite of candidate simplified models identified
    using the R~\cite{R} function dredge from package MuMIn~\cite{MuMIn}. 
    Simplified models were obtained by
    systematically removing all possible combinations of terms from the full model.
    The best-fitting model was then determined to be the model with the fewest terms 
    where $\Delta$AIC\textless2. If several models shared the fewest number of terms 
    and had $\Delta$AIC\textless2, the model with the lowest AIC in that set was chosen as the best-fit
    model.

  \subsection*{Scaling Relationships with S}

@@

    We then simplified each model. To do this, we calculated the AIC of the
    maximal model as well as the AIC's of a suite of candidate simplified
    models identified using the R~\cite{R} function dredge from package
    MuMIn~\cite{MuMIn}.  Simplified models were obtained by systematically
    removing all possible combinations  of terms from the full model with the
    restriction that species richness was retained in all reduced models. The
    best-fitting model was then determined to be the model with the fewest
    terms  where $\Delta$AIC\textless2, as this model is the least likely to suffer from over-fitting.
    As a supplemental check to ensure that variation in sampling effor across
    food webs was not responsible for the trends we observed, we then repeated
    our analyses using jackknifed data sets in which we 1) sequentially removed
    each food web in the dataset and 2) sequentially removed sets of food webs 
    associated with each author of the food webs in the dataset. Parameter 
    estimates varied very little across either series of jackknifes (\emph{Appendix SXX}).
    % If several models shared the fewest
    % number of terms  and had $\Delta$AIC\textless2, the model with the lowest
    % AIC in that set was chosen as the best-fit model.

>>>>

\title{Latitudinal gradients in biotic niche breadth vary across ecosystem types.}
\author{Alyssa R. Cirtwill$^{1,2}$, Daniel B. Stouffer$^{1}$, Tamara N. Romanuk$^{2}$}
\date{\small$^1$Centre for Integrative Ecology\\School of Biological Sciences\\University of Canterbury\\
Private Bag 4800\\Christchurch 8140, New Zealand \\
\medskip$^2$Department of Biology\\
Life Science Centre, Dalhousie University\\1355 Oxford St., P0 BOX 15000\\
Halifax NS, B3H 4R2, Canada\\}

@@

  If the latitude-niche breadth hypothesis is correct, there should also be direct relationships between latitude and the
  degree of specialisation (i.e., the breadth of the Eltonian niche; ~\cite{Elton1927,Leibold2010}) of species within food webs.
  Specifically, narrower niches in the tropics would equate to
  greater specialisation (narrower niches) while constant niche sizes but
  greater productivity would translate to constant specialisation and niche
  width across latitude (Fig.~\ref{concept}).
  Attempts to unravel these effects, however,
  are complicated by known relationships between species richness and many other network properties~\cite{Riede2010}. 
  For example,
  narrower niches imply fewer links per species (i.e., greater specialisation) in the tropics (\cite{Marra1997,Dyer2007};
  but see~\cite{Schleuning2012}). However,
  average numbers of links per species tend to increase in larger food webs~\cite{Dunne2006,Riede2010}. This means
  that a latitudinal effect on specialisation may be obscured by a latitudinal gradient in species richness.

  If this is the case, it may still be possible to uncover effects of latitude
  on specialisation by examining the shape of the scaling  relationship
  between specialisation and species richness over changing latitude. 
  By testing whether latitude affects the scaling of each property with
  species richness, we test for the effects of latitude on specialisation
  implied by the ``latitude-niche breadth hypothesis'' (Fig.~\ref{concept}). 
  If the scaling of specialisation with species richness is weaker in the 
  tropics (i.e., if species gain fewer links, predators, or prey as the size 
  of the network increases), this will indicate narrower niches at the tropics. 
  If, however, the scaling of specialisation with species richness does not 
  vary over latitude, this will indicate that niches are similarly-sized
  worldwide but that there is a broader niche space in the tropics.
  Additionally, as food webs describing different ecosystem types may differ
  in their topology~\cite{Dunne2004,Shurin2006} and may be subject to different
  latitudinal gradients in species richness~\cite{} or productivity~\cite{},
  we also explored the differences in scaling relationships across ecosystem types.
  Here, we use three measures of specialisation; mean links per species, mean
  generality (number of prey), and mean vulnerability (number of predators).

@@

  \subsection*{Gradients over Latitude}

@@

    \noindent where $S_{i}$ is the species richness of web $i$, $L_{i}$ its absolute
    latitude (degrees north or south  regardless of direction), $E_{i}$ is a categorical
    variable indicating the ecosystem type of network $i$ (comprising terms for stream, 
    marine, lake, and terrestrial networks with estuarine
    networks corresponding to $E_{i}=0$) and $\epsilon_{i}$ is a residual error term. 
    We next calculated the AIC
    of the maximal model as well as the AIC's of a suite of candidate simplified models identified
    using the R~\cite{R} function dredge from package MuMIn~\cite{MuMIn}. 
    Simplified models were obtained by
    systematically removing all possible combinations of terms from the full model.
    The best-fitting model was then determined to be the model with the fewest terms 
    where $\Delta$AIC\textless2, as this model is the least likely to suffer from over-fitting.

  \subsection*{Scaling Relationships with Species Richness}

@@

    We then simplified each model. To do this, we calculated the AIC of the
    maximal model as well as the AIC's of a suite of candidate simplified
    models identified using the R~\cite{R} function dredge from package
    MuMIn~\cite{MuMIn}.  Simplified models were obtained by using the same
    procedure as described above. Parameter estimates for the simplified models
    varied very little across either series of jackknifes (\emph{Appendix S3}).
    % If several models shared the fewest
    % number of terms  and had $\Delta$AIC\textless2, the model with the lowest
    % AIC in that set was chosen as the best-fit model.

### 2015-09-02 manuscript/supplement.tex
change roundup
DBS comments

\vspace{0.4 in}
\beginsupplement
\linenumbers
% \begin{landscape}
\subsection*{Appendix S1: Details of food web sources and selection}

@@

\subsection*{Appendix SX: Testing for effects of sampling effort}

@@

    \begin{figure}[h]
    \centerline{\includegraphics*[width=.8\textwidth]{Figures/Jackknife/LS_web.eps}}
    \caption{Estimates of the coefficients in equation XX do not vary significantly
    when jackknifing across food webs. We show the estimates (with 95\% confidence intervals) of each coefficient
    in the equation after the removal of each web in the dataset. 
    The estimate for the full, non-jackknifed
    dataset is indicated by the red, dashed line. All estimates are colour
    coded according the ecosystem type of the removed web.}
    \label{LS_web}
    \end{figure}

    \begin{figure}[h]
    \centerline{\includegraphics*[width=.8\textwidth]{Figures/Jackknife/Gen_web.eps}}
    \caption{Estimates of the coefficients in equation XX do not vary significantly
    when jackknifing across food webs. We show the estimates (with 95\% confidence intervals) of each coefficient
    in the equation after the removal of each web in the dataset. The estimate for the full, non-jackknifed
    dataset is indicated by the red, dashed line. All estimates are colour
    coded according the ecosystem type of the removed web.}
    \label{Gen_web}
    \end{figure}

    \begin{figure}[h]
    \centerline{\includegraphics*[width=.8\textwidth]{Figures/Jackknife/Vul_web.eps}}
    \caption{Estimates of the coefficients in equation XX do not vary significantly
    when jackknifing across food webs. We show the estimates (with 95\% confidence intervals) of each coefficient
    in the equation after the removal of each web in the dataset.  The estimate for the full, non-jackknifed 
    dataset is indicated by the red, dashed line. All estimates are colour
    coded according the ecosystem type of the removed web.}
    \label{Vul_web}
    \end{figure}

    \begin{figure}[h]
    \centerline{\includegraphics*[width=.8\textwidth]{Figures/Jackknife/LS_author.eps}}
    \caption{Estimates of the coefficients in equation XX do not vary significantly
    when jackknifing across authors of the published food webs used in this study. 
    We show the estimates (with 95\% confidence intervals) of each coefficient
    in the equation after the removal of all webs compiled by each author in the dataset.
    The estimate for the full, non-jackknifed dataset is indicated by the red, dashed line.}
    \label{LS_web}
    \end{figure}

    \begin{figure}[h]
    \centerline{\includegraphics*[width=.8\textwidth]{Figures/Jackknife/Gen_author.eps}}
    \caption{Estimates of the coefficients in equation XX do not vary significantly
    when jackknifing across authors of the published food webs used in this study. 
    We show the estimates (with 95\% confidence intervals) of each coefficient
    in the equation after the removal of all webs compiled by each author in the dataset.
    The estimate for the full, non-jackknifed dataset is indicated by the red, dashed line.}
    \label{Gen_web}
    \end{figure}

    \begin{figure}[h]
    \centerline{\includegraphics*[width=.8\textwidth]{Figures/Jackknife/Vul_author.eps}}
    \caption{Estimates of the coefficients in equation XX do not vary significantly
    when jackknifing across authors of the published food webs used in this study. 
    We show the estimates (with 95\% confidence intervals) of each coefficient
    in the equation after the removal of all webs compiled by each author in the dataset.
    The estimate for the full, non-jackknifed dataset is indicated by the red, dashed line.}
    \label{Vul_web}
    \end{figure}

>>>>

\vspace{0.4 in}
\beginsupplement
\linenumbers
% \begin{landscape}

\subsection*{Appendix S1: Details of food web sources and selection}

@@

\subsection*{Appendix S3: Testing for effects of sampling effort}

@@

    \begin{figure}[h]
    \centerline{\includegraphics*[width=.8\textwidth]{Figures/Jackknife/LS_web.eps}}
    \caption{Estimates of the coefficients for scaling of link density with species richness
    in equation 5 (main text) do not vary significantly when jackknifing across food webs. 
    We show the pseudovalues for each coefficient in the equation after the removal of each 
    web in the dataset. The estimate for the full, non-jackknifed dataset is indicated by 
    the red, dashed line.}
    \label{LS_web}
    \end{figure}

    \begin{figure}[h]
    \centerline{\includegraphics*[width=.8\textwidth]{Figures/Jackknife/Gen_web.eps}}
    \caption{Estimates of the coefficients for scaling of generality with species richness
    (see equation 5, main text) do not vary significantly
    when jackknifing across food webs. We show the pseudovalues for each coefficient
    in the equation after the removal of each web in the dataset. The estimate for the full, non-jackknifed
    dataset is indicated by the red, dashed line. }
    \label{Gen_web}
    \end{figure}

    \begin{figure}[h]
    \centerline{\includegraphics*[width=.8\textwidth]{Figures/Jackknife/Vul_web.eps}}
    \caption{Estimates of the coefficients for scaling of vulnerability with species richness
    (see equation 5, main text) do not vary significantly
    when jackknifing across food webs. We show the pseudovalues for each coefficient
    in the equation after the removal of each web in the dataset.  The estimate for the full, non-jackknifed 
    dataset is indicated by the red, dashed line.}
    \label{Vul_web}
    \end{figure}

    \begin{figure}[h]
    \centerline{\includegraphics*[width=.8\textwidth]{Figures/Jackknife/LS_author.eps}}
    \caption{Estimates of the coefficients for scaling of link density with species richness
    in equation 5 (main text) do not vary significantly
    when jackknifing across authors of the published food webs used in this study. 
    We show the pseudovalues for each coefficient
    in the equation after the removal of all webs compiled by each author in the dataset. Authors
    who compiled only a single web were jackknifed as part of the by-web jackknife.
    The estimate for the full, non-jackknifed dataset is indicated by the red, dashed line.}
    \label{LS_web}
    \end{figure}

    \begin{figure}[h]
    \centerline{\includegraphics*[width=.8\textwidth]{Figures/Jackknife/Gen_author.eps}}
    \caption{Estimates of the coefficients for scaling of generality with species richness
    (see equation 5, main text) do not vary significantly
    when jackknifing across authors of the published food webs used in this study. 
    We show the pseudovalues for each coefficient
    in the equation after the removal of all webs compiled by each author in the dataset. Authors
    who compiled only a single web were jackknifed as part of the by-web jackknife.
    The estimate for the full, non-jackknifed dataset is indicated by the red, dashed line.}
    \label{Gen_web}
    \end{figure}

    \begin{figure}[h]
    \centerline{\includegraphics*[width=.8\textwidth]{Figures/Jackknife/Vul_author.eps}}
    \caption{Estimates of the coefficients for the scaling of vulnerability with species richness
    (see equation 5, main text) do not vary significantly
    when jackknifing across authors of the published food webs used in this study. 
    We show the pseudovalues for each coefficient
    in the equation after the removal of all webs compiled by each author in the dataset. Authors
    who compiled only a single web were jackknifed as part of the by-web jackknife.
    The estimate for the full, non-jackknifed dataset is indicated by the red, dashed line.}
    \label{Vul_web}
    \end{figure}
